True versus apparent shapes of bow shocks

### 2017-11-26 papers/Paper1/sec-conic.tex
Finish section on departure function for conics

@@

\begin{figure}
  \includegraphics[width=\linewidth]{figs/conic-departure}
  \caption[]{Parabolic departure function for conic sections.}
  \label{fig:conic-departure}
\end{figure}

@@

The special case of confocal conic sections
(\(\Lambda = \Pi\)) can be written in polar form as
\begin{equation}
  \label{eq:confocal-polar}
  R(\theta) = R_0 \frac {1 + \ecc} {1 + \ecc \cos \theta}
\end{equation}
where \(\ecc = (1 - \Q)^{1/2}\) is the \textit{conic eccentricity}.
For the confocal parabola (\(\ecc = 1\)), the dimensionless reciprocal
radius is therefore \(R_0/R(\theta) = \tfrac12 (1 + \cos \theta)\), which
suggests the following form for a \textit{departure function} that
measures the difference between a given shape \(R(\theta)\) and the
parabola:
\begin{equation}
  \label{eq:departure-function}
  \Depart\bigl(\cos \theta\bigr) = \frac{R_0} {R(\theta)} - \tfrac12 \left( 1 + \cos \theta \right). 
\end{equation}
From equations~\eqref{eq:confocal-polar} and
\eqref{eq:departure-function} it is clear that \(\Depart\) is a linear
function of \(\cos \theta\) for other confocal conics, being positive for
ellipses (\(\ecc < 1\)) and negative for hyperbolae (\(\ecc > 1\)).

>>>>

\begin{figure}
  \includegraphics[width=\linewidth]{figs/conic-departure}
  \caption[]{Parabolic departure function, \(\Depart(\cos\theta)\), for
    conic sections with different planitude and alatude, centered on
    that of the confocal parabola, \((\Pi, \Lambda) = (2, 2)\), which has
    \(\Depart(\cos\theta) = 0\).  Planitude (shown by different line
    colors) determines the slope of \(\Depart\) at the apex,
    \(\cos\theta = 1\).  Alatude (shown by different line widths)
    determines the value of \(\Depart\) at \(\cos\theta = 0\).  All conics
    with \(\Pi = \Lambda\) have departure functions that are straight lines.}
  \label{fig:conic-departure}
\end{figure}

@@

@@

The special case of confocal conic sections
(\(\Lambda = \Pi\)) can be written in polar form as
\begin{equation}
  \label{eq:confocal-polar}
  R(\theta) = R_0 \frac {1 + \ecc} {1 + \ecc \cos \theta}
\end{equation}
where \(\ecc = (1 - \Q)^{1/2}\) is the \textit{conic eccentricity}.
For the confocal parabola (\(\ecc = 1\)), the dimensionless reciprocal
radius is therefore \(R_0/R(\theta) = \tfrac12 (1 + \cos \theta)\), which
suggests the following form for a \textit{departure function} that
measures the difference between a given shape \(R(\theta)\) and the
parabola:
\begin{equation}
  \label{eq:departure-function}
  \Depart\bigl(\cos \theta\bigr) = \frac{R_0} {R(\theta)} - \tfrac12 \left( 1 + \cos \theta \right). 
\end{equation}
From equations~\eqref{eq:confocal-polar} and
\eqref{eq:departure-function} it is clear that \(\Depart\) is a linear
function of \(\cos \theta\) for other confocal conics, being positive for
ellipses (\(\ecc < 1\)) and negative for hyperbolae (\(\ecc > 1\)).
Examples are shown in Figure~\ref{fig:conic-departure} for a grid of 9
conics centered on the confocal parabola, with \((\Pi, \Lambda)\) ranging from
\(3/4\) to \(4/3\) of \((2, 2)\).  The hyperbolae have negative values
of \(\Depart\) in the far wings, with tracks that end at \(\cos \theta_\infty\).

Strictly speaking, the departure function is redundant if one is
interested in only conic sections, since they are fully determined by
\(\Pi\) and \(\Lambda\).  Nonetheless, as we will show in following sections,
it is a useful tool for studying general \(R(\theta)\), being very
sensitive to small variations in the shape.
